Semantic Preserving Bijective Mappings for Representations of Special Functions between Computer Algebra Systems and Document Preparation Systems

### 2018-11-21 ReviewerResponse.txt

We carefully proofread the entire paper again and implemented several minor changes. In the following we response specifically to the reviewers.

Reviewer 1:
As suggested, we made several changes to improve the structure of the paper. In detail we 
•	added subsections in Section 3, 
•	gave an overview of the idea of the translator in Section 4, 
•	removed unnecessary Algorithm 1 in Section 5,
•	added a subsection in Section 5 for a better outline, and 
•	further explained Algorithm 3 (now Algorithm 2).
We also explained the PoM-Tagger, its parsing process, and the lexicon files in more detail.
Further, we added LaTeXML to the related work section and explained briefly why we did not choose MathML as an intermediate data format to perform translations by citing one of our previous publications: Schubotz et al. from 2018 (now included in our references). In this research project, we tested LaTeXML and other state of the art LaTeX to MathML conversion tools. We discovered several problems for those translations. So we decided to work with the PoM-Tagger instead of LaTeXML for our translations.
Also, we did not add further comparisons with LaTeXML to our evaluations section. The reason is that CAS does not understand MathML that contains special functions. Thus we can only compare translations without function calls or implement a solution to overcome this issue for a decent comparison. But this project may legitimate a new publication by itself.

Reviewer 2:
For clarification, we decided to change the wording "Word Processor" to "Document Preparation System.” Since this is part of the title, we did not add that we focus on LaTeX to the title and hope that the Abstract will point that out in an acceptable manner.
Also, instead of creating an Appendix, we decided to add short paragraphs that explain the purpose of the sections. We hope this helps the reader to decide whether he can skip the section, e.g., in case of his lack of interest in Maple “hackings,” or not. 
In the new version of the paper, we implemented all suggested minor changes and added the idea of improved numerical tests by using the mentioned cut region detection to the future work section.

### 2018-11-21 main.tex
Minor changes and added ReviewerResponse

	\Procedure{translate\_macro}{\textit{macro}, \textit{args}, \textit{lexicon}}
	\State \textit{info} = \textit{lexicon}.getInfo(\textit{macro});
	\State \textit{argList} = new List(); \Comment{create a sorted list for the translated arguments.}
	\State \textit{next} = \textit{args}.getNextElement();
	\If{\textit{next} is caret}\label{line:next_caret}
		\State \textit{power} = translateCaret(\textit{next});
		\State \textit{next} = \textit{args}.getNextElement();
	\EndIf

@@

	\State \textit{pattern} = \textit{info}.getTranslationPattern();
	\State \textit{translatedMacro} = \textit{pattern}.fillPlaceHolders(\textit{argList});
	\If{\textit{power} is not \NULL}
		\State \textit{translatedMacro}.add(\textit{power});
	\EndIf
	\State \Return\ \textit{translatedMacro};
	\EndProcedure
	\end{algorithmic}
\end{algorithm}

Algorithm~\ref{alg:macro-translation} explains the \verb|MacroTranslator| without error handling. It has extracted necessary information from the \gls*{pom-pt}, such as how many arguments this function has. In consequence, it also processes the following siblings to translate the arguments. There are some special cases about the next node right after the macro occurs. This node can be
\begin{itemize}
\item an exponent, such as for `\verb|^2|';
\item an optional parameter in square brackets;
\item a parameter in curly brackets (a \textit{sequence} node in the \gls*{pom-pt}) if none of the above;
\item `$@$' symbols if none of the above; or
\item a variable in curly brackets (a \textit{sequence} node) if none of the above.
\end{itemize}

An exponent will be translated and shifted to the end of the translated semantic macro, 
since one usually writes the exponent after the arguments of a function in \gls*{cas}. Hence, the function translates and stores the exponent in Line~\ref{line:next_caret}. One could ask what happens when there is an exponent given before and after the arguments. The \verb|MacroTranslator| only translates the following siblings until each argument is translated. The first exponent will be shifted to the end. If right after the translated macro (with all arguments) follows another exponent, we interpret it as another exponent for the whole previous expression. In that case, it would be the macro with the first translated exponent. Table~\ref{tab:multi-expo} shows an example for the trigonometric cosine function with multiple exponents. As illustrated in Table~\ref{tab:multi-expo}, the translator interprets the input as
\begin{equation}\label{eq:multi-expo}
\left( \cos(x)^n \right)^m.
\end{equation}

\begin{table}[ht]
\centering
\begin{tabular}{lc}
\hline
Displayed As & \rule{0pt}{0.9\normalbaselineskip} $\cos^n(x)^m$ \\
Semantic \LaTeX & \verb|\cos^n@{x}^m| \\
Translated \Maple{} Expression & \verb|((cos(x))^(n))^m|\\
\hline
\end{tabular}
\caption{A trigonometric cosine function example with exponents before and after the argument.}
\label{tab:multi-expo}
\end{table}

>>>>

	\Procedure{translate\_macro}{\textit{macro}, \textit{args}, \textit{lexicon}}
	\State \textit{info} = \textit{lexicon}.getInfo(\textit{macro});\label{line:get_info}
	\State \textit{argList} = new List(); \Comment{create a sorted list for the translated arguments.}
	\State \textit{next} = \textit{args}.getNextElement();
	\If{\textit{next} is caret}\label{line:next_caret}
		\State \textit{power} = translateCaret(\textit{next});
		\State \textit{next} = \textit{args}.getNextElement();
	\EndIf

@@

	\State \textit{pattern} = \textit{info}.getTranslationPattern();
	\State \textit{translatedMacro} = \textit{pattern}.fillPlaceHolders(\textit{argList});\label{line:fill_pattern}
	\If{\textit{power} is not \NULL}\label{line:shifted_exp}
		\State \textit{translatedMacro}.add(\textit{power});
	\EndIf
	\State \Return\ \textit{translatedMacro};
	\EndProcedure
	\end{algorithmic}
\end{algorithm}

The translation process of the \Macro 's is complex, so there is a special class, the \verb|MacroTranslator|, that handles those nodes in the \gls*{pom-pt}. Algorithm~\ref{alg:macro-translation} explains the \verb|MacroTranslator| without error handling. It has extracted necessary information from the \gls*{pom-pt}, such as how many arguments this function has, in Line~\ref{line:get_info}. It also processes the following siblings to translate the arguments. The \verb|MacroTranslator| will be called in Line~\ref{line:transLeaf} in Algorithm~\ref{alg:translation}, since the macro is a leaf node in the \gls*{pom-pt}. The following cases describes the different kinds of the following siblings after a semantic macro node. Those can be
\begin{itemize}
\item an exponent, such as for `\verb|^2|' right after the macro node (Line~\ref{line:next_caret});
\item an optional parameter in square brackets right after the macro node or after an exponent (Line~\ref{line:next_optional});
\item a parameter in curly brackets (a \textit{sequence} node in the \gls*{pom-pt}) if none of the above and no `$@$' symbols were passed yet (Line~\ref{line:trans_paras});
\item `$@$' symbols (Line~\ref{line:skip_ats}); or
\item a variable in curly brackets (a \textit{sequence} node) after the `$@$' symbols were passed. (Line~\ref{line:trans_vars})
\end{itemize}

All cases before the `$@$' symbols are optional. The \verb|MacroTranslator| removes all following siblings according to the number of expected parameters and variables. Parameter and variable nodes are translated separately. If an exponent was registered right after the semantic macro node, it will be shifted to the end in Line~\ref{line:shifted_exp}. The macro itself will be translated by putting all translated parameters and variables into the translation pattern (Line~\ref{line:fill_pattern}).

Following siblings after the macro was translated (with all arguments) do not belong to the semantic macro. If the next node is an exponent the translated macro is the base. Table~\ref{tab:multi-expo} shows an example for the translation of the trigonometric cosine function with multiple exponents.

\begin{table}[ht]
\centering
\begin{tabular}{lccc}
 & Semantic \LaTeX & & \Maple{} \\
\hline
Text Representation & \rule{0pt}{1.0\normalbaselineskip} \verb|\cos^n@{x}^m| & $\overset{\langMaple}{\mapsto}$ & \verb|((cos(x))^(n))^m| \\
Displayed As & \rule{0pt}{1.0\normalbaselineskip} $\cos^n(x)^m$ & & $\left( \cos(x)^n \right)^m$ \\
\hline
\end{tabular}
\caption{A trigonometric cosine function example with exponents before and after the argument.}
\label{tab:multi-expo}
\end{table}
